---
title: "All steps to delineate Level 4 units based on marine provinces (MEOW)"
output: html_document
editor_options:
  chunk_output_type: console
---

First load all libraries and set-up project environment variables.

```{r}
#!/usr/bin/R --vanilla
require(dplyr)
require(magrittr)
require(readr)
require(sf)
require(units)
require(leaflet)
require(tidyr)

source(sprintf("%s/proyectos/IUCN-GET/MFT1.2-intertidal-forest/env/project-env.R",Sys.getenv("HOME")))
setwd(work.dir)

```

Read spatial data of ecoregions and species distribution data:

```{r}
meow <- read_sf("mangrove-type-data-sources.vrt","meow")

##(load(sprintf("%s/www/Rdata/mgv-species-occurrence.rda",script.dir)))
(load(sprintf("%s/www/Rdata/species-occurrence.rda",script.dir)))

meow %>% left_join({mprov.xy %>% st_drop_geometry %>% transmute(ECO_CODE, n_key_mgv_spp=n_mgv_sppp, n_add_mgv_spp=rowSums(mtzXprov>0))},by="ECO_CODE") -> meow_mgv_summary

```

We intersect the marine ecoregions with different global maps of mangrove distribution:

```{r}
if (!exists("usgs.meow.its")) {
  usgs.meow.its <- read_sf("intersection-usgs-meow-provs.gpkg")
  usgs.meow.its %<>% mutate(area=st_area(geometry))
}
meow_mgv_summary %<>% left_join({usgs.meow.its %>% st_drop_geometry %>% group_by(ECO_CODE) %>% summarise(n_usgs_2000=n(),area_usgs_2000=sum(area) %>% set_units('km^2'))},by="ECO_CODE")

if (!exists("gmw.meow.1996")) {
  gmw.meow.1996 <- read_sf("intersection-gmw-meow-provs-1996.gpkg")
  gmw.meow.1996 %<>% mutate(area=st_area(geometry))
}
meow_mgv_summary %<>% left_join({gmw.meow.1996 %>% st_drop_geometry %>% group_by(ECO_CODE) %>% summarise(n_gmw_1996=n(),area_gmw_1996=sum(area) %>% set_units('km^2'))},by="ECO_CODE")

if (!exists("gmw.meow.2016")) {
  gmw.meow.2016 <- read_sf("intersection-gmw-meow-provs-2016.gpkg")
  gmw.meow.2016 %<>% mutate(area=st_area(geometry))
}
meow_mgv_summary %<>% left_join({gmw.meow.2016 %>% st_drop_geometry %>% group_by(ECO_CODE) %>% summarise(n_gmw_2016=n(),area_gmw_2016=sum(area) %>% set_units('km^2'))},by="ECO_CODE")

if (!exists("mgt.meow.2016")) {
  mgt.meow.2016 <- read_sf("intersection-type-meow-provs-2016.gpkg")
  mgt.meow.2016 %>% mutate(area=st_area(geometry)) %>%  st_centroid -> mgt_points
}

```

Summaries of number of species and area of mangrove formation (ecosystem) detected per marine ecoregion:

```{r}
meow_mgv_summary %<>% rowwise() %>%
  mutate(min_area = min(c_across(starts_with("area"))),max_area = max(c_across(starts_with("area")))) %>% ungroup %>% st_sf(sf_column_name="geom")

meow_mgv_summary %<>% mutate(native=if_else(is.na(n_key_mgv_spp) | PROV_CODE %in% c(37:40),FALSE,TRUE))
meow_mgv_summary %<>% mutate(mgv_eco=case_when(
  !is.na(area_gmw_2016)~'recently detected',
  !is.na(area_usgs_2000)~'previously detected',
  !is.na(area_gmw_1996)~'previously detected',
  ECOREGION %in% c("Leeuwin","Lord Howe and Norfolk Islands") ~'known but unmapped',
  ECO_CODE %in% c(20106,20200,20201)~'absent',
  PROV_CODE %in% c(4,5,8,50,62)~'absent',
  native~'unmapped',
  TRUE~'absent'))

meow_mgv_summary %>% st_drop_geometry%>% dplyr::select(mgv_eco,native) %>% table

meow_mgv_summary %>% filter(!mgv_eco %in% 'absent') %>% summarise(total=n_distinct(PROV_CODE))
meow_mgv_summary %>% filter(!mgv_eco %in% c('absent','unmapped')) %>% summarise(total=n_distinct(PROV_CODE))

meow_mgv_summary %<>% mutate(step_1=case_when(
  (mgv_eco %in% 'absent')~'absent',
  ECOREGION %in% c('Maldives','Leeuwin','Lord Howe and Norfolk Islands')~'Included (unmapped, but known)',
  (!native)~'included (non-native)',
  (mgv_eco %in% 'unmapped')~ 'Consider exclusion',
  (PROV_CODE %in% c(3))~ 'Consider exclusion',
  PROV_CODE %in% c(6,9,11,47,53,54,16,28,44,45) ~ "Potential merge",
  PROV_CODE %in% c(38,39,40) ~ "Potential merge (non-native)",
  PROV_CODE %in% c(12,18,21) ~ "Potential split",
  ECOREGION %in% c("Lord Howe and Norfolk Islands") ~'Potential merge (unmapped, but known)',
  PROV_CODE %in% c(56,57,58,13,51) ~ "Potential modification",
  TRUE ~ "Included"))

meow_mgv_summary %>% filter(PROV_CODE %in% c(36)) %>% dplyr::select(ECOREGION,n_key_mgv_spp,n_add_mgv_spp,min_area,max_area,native,mgv_eco)
meow_mgv_summary %>% filter(PROV_CODE %in% c(54)) %>% dplyr::select(ECOREGION,n_key_mgv_spp,n_add_mgv_spp,min_area,max_area,native,mgv_eco)


```
|PROV CODE| PROVINCE | exclusion |
|---|---|---|
|45| Warm Temp SE Pacific| keep one ecoregion|
|4| Mediterranean| exclude|
|8| Cold Temp NW Pacific | exclude|
|5| Cold Temp NW Atlantic| exclude|
|62| Benguela| exclude|
|22| C Indian Ocean Islands | exclude eco code 20106|
|36| Lord Howe and Norfolk Islands |Include!|


Create dataset with selected provinces and data to display:
```{r}
pre_units <- meow_mgv_summary  %>% filter(!step_1 %in% 'absent') %>% group_by(PROV_CODE,PROVINCE,step_1) %>% summarise(area_2000=sum(area_usgs_2000,na.rm=T), area_2016=sum(area_gmw_2016,na.rm=T))

mprov.xy %>% st_drop_geometry %>% group_by(PROV_CODE,PROVINCE) %>% summarise(across(Dolichandrone.spathacea:Ceriops.australis, max)) %>% rowwise() %>% transmute(PROV_CODE,PROVINCE,sum=sum(c_across(Dolichandrone.spathacea:Ceriops.australis))) %>% ungroup -> mtz

pre_units %<>% left_join(mtz,by=c('PROV_CODE','PROVINCE'))  %>% transmute(PROVINCE,`Mangrove Area in 2000 (km^2)`=round(area_2000,2),`Mangrove Area in 2016 (km^2)`=round(area_2016,2),`Mangrove species`=sum,Comment=step_1)
save(file=sprintf("%s/www/Rdata/selected-units.rda",script.dir),pre_units)

```


Preview maps before any modifications with leaflet

```{r}
m <- leaflet(options = leafletOptions(worldCopyJump = TRUE,zoomControl = TRUE))  %>%  addProviderTiles(providers$Esri.OceanBasemap) %>%
  setView(lng = 0, lat = 0, zoom = 2)

clrs <- colorFactor("Dark2",domain=pre_units$Comment)

labels <- sprintf("%s<br/>%s<br/>%s", pre_units$PROV_CODE, pre_units$PROVINCE, pre_units$Comment) %>% lapply(htmltools::HTML)

m %>% addPolygons(data = pre_units, fillColor = ~clrs(Comment), highlightOptions = highlightOptions(weight = 2, color = 'black'), color = 'grey', weight = 0.4, fillOpacity = 0.45,label=labels)


```

All other steps

```{r}

meow_mgv_summary %<>% mutate(step_2=case_when(
  PROV_CODE %in% c(45,53,54,36,38,39,40) ~ "Merge",
  PROV_CODE %in% c(11,6,47,9,28,16,51) ~ "Do not merge",
  TRUE ~ as.character(NA)))

meow_mgv_summary %<>% mutate(step_3=case_when(
  PROV_CODE %in% c(12) ~ "Split",
  PROV_CODE %in% c(18,21) ~ "Do not split",
  TRUE ~ as.character(NA)))

meow_mgv_summary %<>% mutate(step_4=case_when(
  PROV_CODE %in% c(56,57,58) ~ "Modify",
  PROV_CODE %in% c(51,13) ~ "Do not modify",
  TRUE ~ as.character(NA)))

meow_mgv_summary %<>% mutate(description=case_when(
  ECO_CODE %in% c(20173,20174,20172) ~ "Kept as part of distinct unit",
  PROV_CODE %in% c(11,6,16,9) ~ "Boundary between tropical and temperate",
  ECO_CODE %in% c(20175,20209,20206,20207,20208,20211,20210,20209,20197,25196,25199) ~ "Merged to keep continuity in distribution",
  ECO_CODE %in% c(20169,20164,20181,20074,20192,20097,20099,20093,20120,20114,20150) ~ "Kept as part of province, occurrence not confirmed",
  ECO_CODE %in% c(20180,20192) ~ "Boundary between tropical and temperate",
  PROV_CODE %in% c(13) ~ "Keeping original boundaries of the province",
  PROV_CODE %in% c(28) ~ "Keeping separate as ditinct unit",

  ECO_CODE %in% c(20062) ~ "Split from rest of province, isolated occurrence",
  PROV_CODE %in% c(12,17,20,19,18,23,24,27,26,25,30,29,31:35,55,56,58) ~ "Kept as part of distinct unit",
  ECO_CODE %in% c(20105,20151) ~ "Unmapped but known occurrence",
  PROV_CODE %in% c(21) ~ "Might be assessed as two units",
  PROV_CODE %in% c(37:40) ~ "Introduced",
  TRUE ~ "")
  )

```


Nomenclature:

```{r}

prefix <- "Intertidal forests and shrublands"
prefix <- "Mangroves"
code_format <- "MFT1.2_4_%s_%02d%s"

meow_mgv_summary %<>% mutate(
  unit_code = case_when(
    ECOREGION %in% "Bermuda" ~ sprintf(code_format,"MP",PROV_CODE,"b"),
    PROV_CODE %in% 12 ~ sprintf(code_format,"MP",PROV_CODE,"a"),
    ECO_CODE %in% 20104 ~ sprintf(code_format,"MP",PROV_CODE,"b"),
    ECO_CODE %in% 20103 ~ sprintf(code_format,"MP",PROV_CODE,"a"),
    PROV_CODE %in% c(43,45) ~ sprintf(code_format,"MP",43,"x"),
    PROV_CODE %in% c(53,54) ~ sprintf(code_format,"MP",53,"m"),
    ECOREGION %in% "Leeuwin" | PROVINCE %in% "West Central Australian Shelf" ~ sprintf(code_format,"MP",58,"x"),
    ECOREGION %in% c("Great Australian Bight", "South Australian Gulfs", "Western Bassian")~ sprintf(code_format,"MP",57,"m"),
    PROV_CODE %in% c(56) ~ sprintf(code_format,"MP",56,"r"),
    PROV_CODE %in% c(36,55) ~ sprintf(code_format,"MP",55,"x"),
    PROV_CODE %in% c(38,39,40) ~ sprintf(code_format,"MP",38,"x"),
    TRUE ~ sprintf(code_format,"MP",PROV_CODE,"")),
  unit_name = case_when(
  step_1 %in% 'absent' | PROV_CODE %in% c(3) ~ 'absent',
  ECO_CODE %in% c(20106,25195,20200,20201,25198)~'absent',
  ECOREGION %in% "Bermuda" ~ sprintf("%s of %s",prefix,"Bermuda"),
  ECO_CODE %in% 20104 ~ sprintf("%s of %s",prefix,"South India and Sri Lanka"),
  ECO_CODE %in% 20103 ~ sprintf("%s of %s",prefix,"Western India"),
  PROV_CODE %in% c(43,45) ~ sprintf("%s of %s",prefix,"East and Southeastern Pacific Coast"),
  PROV_CODE %in% c(53,54) ~ sprintf("%s of %s",prefix,"New Zealand"),
  PROV_CODE %in% c(36,55) ~ sprintf("%s of %s",prefix,"East Central Australian Shelf"),
  ECOREGION %in% "Leeuwin" | PROVINCE %in% "West Central Australian Shelf" ~ sprintf("%s of %s",prefix,"Western Australian Shelf and Leeuwin"),
  ECOREGION %in% c("Great Australian Bight", "South Australian Gulfs", "Western Bassian")~ sprintf("%s of %s",prefix,"South Australia"),
  PROV_CODE %in% c(38,39,40) ~ sprintf("%s of %s",prefix,'the Central Pacific'),
  TRUE ~ sprintf("%s of %s",prefix,PROVINCE))
)

meow_mgv_summary %>% st_drop_geometry %>% filter(unit_name != 'absent') %>% distinct(unit_code,unit_name,native) %>% arrange(unit_code) %>% print.AsIs()
```


```{r}
meow_mgv_summary
post_units <- meow_mgv_summary  %>% filter(!unit_name %in% 'absent') %>% group_by(unit_code,unit_name,native) %>% summarise

post_units_components <- meow_mgv_summary %>% st_drop_geometry %>% filter(!unit_name %in% 'absent')
#post_units %<>% left_join({meow_mgv_summary %>% st_drop_geometry %>% filter(!unit_name %in% 'absent') %>% group_by(unit_name,PROVINCE) %>% summarise(ecoregions=paste0(ECOREGION,collapse=", ")) %>% group_by(unit_name) %>% summarise(description=paste(sprintf("Ecoregions %s of the **<emph>%s</emph>** province",ecoregions,PROVINCE),collapse="; "))},by="unit_name")
save(file=sprintf("%s/www/Rdata/selected-units.rda",script.dir), pre_units, post_units, post_units_components)

```


Visualise this

```{r}
clrs <- colorFactor("Dark2",domain=post_units$native)

labels <- sprintf("%s <strong>%s</strong><br/>Native:%s", post_units$unit_code,post_units$unit_name, post_units$native) %>% lapply(htmltools::HTML)

m %>% addPolygons(data = post_units, fillColor = ~clrs(native), highlightOptions = highlightOptions(weight = 2, color = 'black'), color = 'grey', weight = 0.4, fillOpacity = 0.45,label=labels)

```


We will include point layers with the location of polygons


```{r}
gmw.2016.xy <- gmw.meow.2016 %>% st_centroid()
usgs.2000.xy <- usgs.meow.its %>% st_centroid()

gmw.2016.xy %<>% mutate(area=area %>% set_units(km^2))
usgs.2000.xy %<>% mutate(area=area %>% set_units(km^2))

# too many points for a global map
#m %>%  addCircleMarkers(data=gmw.2016.xy,radius=sqrt(gmw.2016.xy$area),color = 'green', fillColor = 'green', opacity = .3,group="Mangrove occurrences")

mgt_points %<>% left_join({meow_mgv_summary %>% st_drop_geometry() %>% select(ECO_CODE,unit_code,unit_name)},by="ECO_CODE")

# mgt_points %>% st_drop_geometry %>% group_by(unit_name,Class,Sedimentar) %>% summarise(total=sum(area) %>% drop_units) %>% pivot_wider(id=c(unit_name), names_from=c(Class,Sedimentar), values_from=total) %>% replace(is.na(.), 0) -> units_type_areas

mgt_points %>% st_drop_geometry %>% group_by(unit_code,unit_name,Class,Sedimentar) %>% summarise(npols=n(),total=sum(area) %>% set_units('km^2')) -> units_table

save(file=sprintf("%s/www/Rdata/mgt-point-data.rda",script.dir), mgt_points, units_table)


```


```{r}
# unzip(sprintf("%s/admin/global/World-Bank/wb_countries_admin0_10m.zip",gis.data))
adm0 <- read_sf("WB_countries_Admin0_10m")
post_units_countries <- post_units %>% select(unit_code) %>% st_intersection(adm0 %>% filter(!ISO_A2 %in% -99) %>% select(WB_NAME, TYPE, ISO_A2, CONTINENT, REGION_UN, SUBREGION, REGION_WB)) %>% st_drop_geometry()

# Alternative shortnames:
post_units %<>% 
  transmute(unit_code,unit_name,shortname=gsub(
    "Mangroves of |the |orth|ast|est|ern|outh|ical|erate|entral", "", unit_name) %>% 
      gsub("([NS])w ","\\1W ",.) %>% 
      gsub("([NS])e ","\\1E ",.) %>% 
      gsub("E C ","EC ",.) %>% 
      gsub(" and ","+",.)
    ,native) 
## post_units %>% st_drop_geometry() %>% select(unit_name,shortname) %>% print.AsIs()


save(file=sprintf("%s/www/Rdata/selected-units.rda",script.dir), pre_units, post_units, post_units_components, post_units_countries)

```

Finally we will add these to the database:

```{r}
#!/usr/bin/R --vanilla
require(dplyr)
require(magrittr)
require(readr)
require(RPostgreSQL)
require(sf)

source(sprintf("%s/proyectos/IUCN-GET/MFT1.2-intertidal-forest/env/project-env.R",Sys.getenv("HOME")))
setwd(work.dir)

drv <- dbDriver("PostgreSQL") ## remember to update .pgpass file

con <- dbConnect(drv, dbname = dbinfo['database'],
              host = dbinfo['host'],
              port = dbinfo['port'],
              user = dbinfo['user'])

mgt.meow.its <- read_sf("intersection-type-meow-provs.gpkg")

mgt.meow.its %>% st_drop_geometry -> mgt.table

mgt.table %>% distinct(Class,Sedimentar,PROV_CODE,PROVINCE) %>% mutate(GT=as.numeric(as.factor(Class))*10 + as.numeric(as.factor(Sedimentar))) -> smr.table

smr.table %>% transmute(qry=
  str_glue(
    "INSERT INTO  bio_geo_ecotype (bge_code,code,name,shortdesc,status,contributors)
    VALUES(
      'MFT1.2-MP{PROV_CODE}-GT{GT}','MFT1.2','MFT1.2 Intertidal forests and shrublands of biophysical type *{Class} ({Sedimentar})* in the *{PROVINCE}* province', 'Spatial overlap of Global biophysical typology of mangroves (class={Class} and sedimentar={Sedimentar}) and the marine province of {PROVINCE} (PROV_CODE={PROV_CODE}).','draft','{{JRFP,Global Mangrove Alliance}}'
    ) ON CONFLICT DO NOTHING")) %>% pull-> qries

for (qry in qries) {
  dbSendQuery(con,qry)
}

mgt.meow.its %>% group_by(Class,Sedimentar,PROV_CODE) %>% summarise(area=sum(st_area(geometry))) -> mgt.area
```


### References

Ellison, J. C. (1996). Pollen Evidence of Late Holocene Mangrove Development in Bermuda. Global Ecology and Biogeography Letters, 5(6), 315–326. https://doi.org/10.2307/2997587