---
title: "Analyse species distribution data per province "
editor_options:
  chunk_output_type: console
---

Load libraries and set-up working environment:

```{r}
require(dplyr)
require(magrittr)
require(sf)
require(units)
require(vegan)

source(sprintf("%s/proyectos/IUCN-GET/MFT1.2-intertidal-forest/env/project-env.R",Sys.getenv("HOME")))
setwd(work.dir)
```

Here we will intersect the polygon of _Marine provinces_ from [MEOW](https://www.worldwildlife.org/publications/marine-ecoregions-of-the-world-a-bioregionalization-of-coastal-and-shelf-areas) with species distribution ranges from the RLTS database and records from GBIF.

Range distribution maps from key mangrove species was provided by the RLTS team in a shapefile.

```{r}
mangroves <- read_sf(sprintf("%s/species-dist/global/IUCN_RLTS/MANGROVES.shp",gis.data))
```

For other species associated with mangroves we ran a script to download GBIF data in the background.

```sh
cd $WORKDIR
nohup Rscript $SCRIPTDIR/inc/R/download-from-GBIF.R > nohup-gbif-qries.out &
```

We then use rsync to update the results in the shared drive

```sh
export srcdir=$GISDATA/species-dist/global/GBIF/
export dstdir=$zID@kdm.restech.unsw.edu.au:/srv/scratch/cesdata/gisdata/species-dist/global/GBIF
# this is a dry run: remove -n flag for actual syncing
rsync -gloptrunv $srcdir/* $dstdir
```

We transformed these distribution maps and records into a matrix of species per marine ecoregion:

```sh
cd $WORKDIR
nohup Rscript $SCRIPTDIR/inc/R/spp-per-province.R > nohup-spp-x-province.out &
```

# Biotic dissimilarity and cluster analysis

Now we can use these matrices to calculate biotic dissimilarity between ecoregions and provinces. First for key mangrove species

```{r}
(load(sprintf("%s/www/Rdata/mgv-species-occurrence.rda",script.dir)))
(load(sprintf("%s/www/Rdata/species-occurrence.rda",script.dir)))

mgv.spp <- prov.mgv.spp %>% st_drop_geometry %>% select(Dolichandrone.spathacea:Ceriops.australis)
mga.spp <- mtzXprov[,(gsub("_",".",colnames(mtzXprov)) %in% colnames(mgv.spp))]

ss <- rowSums(mga.spp)>0

## this looks like a good combination (homogeneous groupings)
d.mgv <- vegdist(mgv.spp[ss,],'kulczynski',binary=T)
d.mga <- vegdist(mga.spp[ss,],'kulczynski',binary=T)

mantel(d.mgv,d.mga)


h.mgv <- hclust(d.mgv,method='ward.D2')
h.mga <- hclust(d.mga,method='ward.D2')

plot(h.mgv)
plot(h.mga)

```

The larger scale groups are consistent with provinces, except for these

```{r}
hc <- as.dendrogram(h.mgv)
tt <- table(prov.mgv.spp$PROVINCE,cutree(h.mgv,k=6))
table(rowSums(tt>0))


hc <- as.dendrogram(h.mga)
tt <- table(prov.mgv.spp$PROVINCE,cutree(h.mga,k=6))
table(rowSums(tt>0))


labelColors = c("#CDB380", "#036564", "#EB6841", "#EDC951")

# ls <- meow[match(h1$labels,meow$ECO_CODE),c("PROVINCE","ECOREGION")]
#clusLabel = if_else(ls$PROVINCE %in% rownames(tt)[rowSums(tt>0)>1],ls$PROVINCE,"o")
clusMember = cutree(h1, 6)

# function to get color labels
colLab <- function(n) {
  if (is.leaf(n)) {
    a <- attributes(n)
    qry <- meow %>% st_drop_geometry %>% filter(ECO_CODE %in% a$label) %>% pull(PROVINCE)
    newLabel <- if_else(qry %in% rownames(tt)[rowSums(tt>0)>1],qry,"o")
    labCol <- labelColors[clusMember[which(names(clusMember) == a$label)]]
    attr(n, "nodePar") <- c(a$nodePar, lab.col = labCol)
    attr(n,"label")  <- newLabel
  }
  n
}

# using dendrapply
clusDendro = dendrapply(hc, colLab)

# make plot
plot(clusDendro, main = "Cool Dendrogram")

```


```{r}
mgt.2016.pols %>%  pull(PROV_CODE) %>% unique -> slc
mgt.2016.pols %>%  pull(ECO_CODE) %>% unique -> ecos

#meow %>% filter(PROV_CODE %in% slc) %>% group_by(PROV_CODE,PROVINCE) %>% summarise(geom=st_union(geom)) -> mprovs

#unzip(sprintf("%s/admin/global/World-Bank/wb_land_10m.zip",gis.data))
#unzip(sprintf("%s/admin/global/World-Bank/wb_boundaries_geojson_lowres.zip",gis.data))
#wbland <- read_sf("WB_Land_10m/WB_Land_10m.shp")
wbland <- read_sf("WB_Boundaries_GeoJSON_lowres/WB_Coastlines_10m_lowres.geojson")

#meow <- read_sf("mangrove-type-data-sources.vrt","meow")
#meow %>% filter(PROV_CODE %in% slc) %>% group_by(PROV_CODE,PROVINCE) %>% summarise(geom=st_union(geom)) -> mprovs
```

# Mangrove species

Data on mangroves was requested from the RLTS team and the file downloaded.

```{r}

mangroves %>% st_drop_geometry %>% group_by(kingdom,phylum,class,order_) %>% summarise(nspp=n_distinct(id_no))
mangroves %>% st_make_valid %>% st_intersects(post_units,sparse=F) -> tst

gsub("Mangroves of ", "",colnames(tst))
colnames(tst) <- post_units %>% st_drop_geometry %>% transmute(nms=gsub("Mangroves of ", "",unit_name)) %>% pull

# Kingdom plantae, phylum TRACHEOPHYTA 
mangroves %>% st_drop_geometry %>% select(id_no,binomial,legend,class,order_,family,code) %>% bind_cols(tibble(data.frame(tst))) -> key_spp_data

write_csv(key_spp_data,file=sprintf("%s/mangrove-species-table.csv",work.dir))


mangrove_species %>% filter(!(taxonid %in% key_spp_data$id_no) & majorimportance %in% "Yes" & suitability %in% 'Suitable') %>% distinct(taxonid,.keep_all = TRUE) %>%  select(taxonid,kingdom_name,phylum_name,class_name,order_name,family_name,genus_name,scientific_name,taxonomic_authority,category,main_common_name) -> mga_spp

mga_spp %<>% mutate(mcolumn=match(scientific_name,gsub("_"," ",colnames(mtzXprov))))

mga_spp %>% filter(is.na(mcolumn))


mangrove_species %>% filter(majorimportance %in% "Yes" & suitability %in% 'Suitable') %>% summarise(total=n(),nspp=n_distinct(taxonid))  
mangrove_species %>% select(taxonid,code,suitability,season,majorimportance) -> spp_habitat

require(tidyr)

spp_habitat %>% pivot_wider(taxonid,names_from=code,values_from=c(suitability,season,majorimportance))

spp_habitat %>% filter(majorimportance %in% "Yes" & suitability %in% 'Suitable') %>% select(majorimportance,season) %>% table()


load(sprintf("%s/www/prov-data.rda",script.dir))

load()
gsub("\\.","_",colnames(prov.mgv.spp)) %in% colnames(sppXprov)

```

Measure biotic simmilarity between provinces:




# All (other) species



```{r}

load(sprintf("%s/www/prov-data.rda",script.dir))


wbland %>% st_transform(crs=st_crs(mgt.2016.pols)) %>% st_simplify(dTolerance=4) -> world.xy
mi.rda <- sprintf("%s/www/Rdata/gis.rda",script.dir)
save(file=mi.rda,world.xy)

```

We create a equal area grid on top of marine provinces.


Measure biotic simmilarity between provinces:

```{r}
ss <- rowSums(mtzXprov)>0

## this looks like a good combination (homogeneous groupings)
d1 <- vegdist(sppXprov[ss,],'kulczynski',binary=T)
h1 <- hclust(d1,method='ward.D2')

lbls <- gsub("West|Western|west|western","W",mprov.xy$PROVINCE)
lbls %<>% gsub("East|Eastern|east|eastern","E",.)
lbls %<>% gsub("North|Northern","N",.)
lbls %<>% gsub("South|Southern","S",.)
lbls %<>% gsub("Central","C",.)
lbls %<>% gsub("Tropical","Trop",.)
lbls %<>% gsub("Temperate","Temp",.)


plot(h1,labels=lbls)
table(cutree(h1,h=.5))
abline(h=.5,lty=2,col=2)

mprov.xy$k_group <- cutree(h1,h=.5)
mprov.xy %>%  group_by(k_group) %>% summarise -> tst
plot(tst)
```



Cluster analysis for grid cells

```{r}
ss <- rowSums(sppXcell)>0

## this looks like a good combination (homogeneous groupings)
d1 <- vegdist(sppXcell[ss,],'kulczynski',binary=T)
h1 <- hclust(d1,method='ward.D2')
plot(h1)
table(cutree(h1,k=35))

spp.grid %>% filter(ss) %>% transmute(k_group=cutree(h1,k=35)) %>% plot
spp.grid %>% filter(ss) %>% transmute(k_group=cutree(h1,k=35)) %>% group_by(k_group) %>% summarise -> tst
plot(tst)
```


```{r}
head(res)
table(res$kingdom_name,res$habitat)

res %>% group_by(class_name) %>% summarise(total=n_distinct(scientific_name))

res %>% group_by(order_name) %>% summarise(total=n_distinct(scientific_name))
res %>% group_by(order_name) %>% summarise(total=n_distinct(scientific_name)) %>% arrange(desc(total))
```

Check first how many records are available in GBIF for each species:

```{r}
summary(res$gbif_records)
summary(res$gbif_xy_records)
```
