---
title: "Analyse species distribution data per province "
editor_options:
  chunk_output_type: console
---


We intersected the polygon of _Marine provinces_ from [MEOW](https://www.worldwildlife.org/publications/marine-ecoregions-of-the-world-a-bioregionalization-of-coastal-and-shelf-areas) and species distribution records from GBIF

We run a script to download GBIF data in the background.

```sh
cd $WORKDIR
nohup Rscript $SCRIPTDIR/inc/R/download-from-GBIF.R > nohup-gbif-qries.out &
```

We then use rsync to update the results in the shared drive

```sh
export srcdir=$GISDATA/species-dist/global/GBIF/
export dstdir=$zID@kdm.restech.unsw.edu.au:/srv/scratch/cesdata/gisdata/species-dist/global/GBIF
# this is a dry run: remove -n flag for actual syncing
rsync -gloptrunv $srcdir/* $dstdir 
```


```{r}
require(dplyr)
require(magrittr)
require(sf)
require(units)

source(sprintf("%s/proyectos/IUCN-GET/MFT1.2-intertidal-forest/env/project-env.R",Sys.getenv("HOME")))
setwd(work.dir)

mgt.2016.pols <- read_sf("eck4-mangrove-type-provs.gpkg")
mgt.2016.pols %>%  pull(PROV_CODE) %>% unique -> slc

#unzip(sprintf("%s/admin/global/World-Bank/wb_land_10m.zip",gis.data)) 
wbland <- read_sf("WB_Land_10m/WB_Land_10m.shp")

#meow <- read_sf("mangrove-type-data-sources.vrt","meow")
#meow %>% filter(PROV_CODE %in% slc) %>% group_by(PROV_CODE,PROVINCE) %>% summarise(geom=st_union(geom)) -> mprovs
load(sprintf("%s/www/prov-data.rda",script.dir))

## st_wrap_dateline to solve the international date line problem:
# mprovs  %>% st_wrap_dateline %>% plot
# mprovs %>% st_wrap_dateline(c("WRAPDATELINE=YES", "DATELINEOFFSET=50")) %>% st_geometry %>% plot

## filter provinces with data in 2016 + Lusitania
mprov.xy <- prov_summary %>% filter(PROV_CODE %in% c(3,slc)) %>% st_wrap_dateline() %>% st_transform(crs=st_crs(mgt.2016.pols))
plot(st_geometry(mprov.xy))

wbland %>% st_transform(crs=st_crs(mgt.2016.pols)) -> world.xy
mi.rda <- sprintf("%s/www/gis.rda",script.dir)
save(file=mi.rda,world.xy)

```

We create a equal area grid on top of marine provinces. 

```{r}
raw.grid <- st_make_grid(mprov.xy,cellsize=500000)

spp.grid <- st_as_sf(tibble(code=1:length(raw.grid)),raw.grid)
spp.grid %>% plot
plot(st_geometry(spp.grid))
plot(mprov.xy['PROV_CODE'],add=T)
```


```{r}
archs <- list.files(sprintf("%s/species-dist/global/GBIF/%s",gis.data,projectname),recursive=T,pattern='.rda$',full.names=T)

mi.rda <- sprintf("%s/www/species-occurrence.rda",script.dir)
if (file.exists(mi.rda)) {
  load(mi.rda)
} else {
  load(grep("overview",archs,value=T))
  sppXprov <- tibble(.rows=nrow(mprov.xy))
  sppXcell <- tibble(.rows=length(raw.grid))
}

for (arch in grep("occ-data-spp",archs,value=T)) {
  objs <- (load(arch))  
  for (oo in objs) {
    gbif.data <- get(oo)
    for (spp in unique(names(gbif.data$gbif$data))) {
      if (!(spp %in% colnames(sppXprov)) & !(spp %in% colnames(sppXcell)) & (nrow(gbif.data$gbif$data[[spp]])>0)) {
        xys <- gbif.data$gbif$data[[spp]] %>% dplyr::select(name,latitude,longitude,key) %>% filter(!is.na(longitude)& !is.na(latitude))
        if (nrow(xys)>0) {
          spps <- st_as_sf(xys,coords=c("longitude","latitude"),crs=st_crs(meow)) %>% st_transform(st_crs(mgt.2016.pols))
          mat <- spps %>% st_intersects(mprov.xy,sparse=F) 
          sppXprov %<>% bind_cols(tibble(colSums(mat),.name_repair=~c(spp)))
          mat <- spps %>% st_intersects(spp.grid,sparse=F) 
          sppXcell %<>% bind_cols(tibble(colSums(mat),.name_repair=~c(spp)))
          save(file=mi.rda,sppXcell,sppXprov,mangrove_species,mprov.xy)
        }
      }
    }
  }
}
#
# not sure why it produces duplicated names: but they seem to include the same records...
#names(spps.lt1500$gbif$data)[duplicated(names(spps.lt1500$gbif$data))]
#grep("Lycaon_pictus",names(spps.lt1500$gbif$data))

dim(sppXcell)
dim(sppXprov)
```

For species with very large number of records, we have to summarise data with a different approach
```{r}
grdXprov <- tibble(.rows=nrow(mprov.xy))
grdXcell <- tibble(.rows=length(raw.grid))
list_spps <- c()
for (arch in grep("occ-data-grid",archs,value=T)) {
  objs <- try(load(arch))  
  if (any(class(objs)=='try-error')) {
    cat(sprintf('error with %s!\n',basename(arch)))
    next
  }
  for (oo in objs) {
    gbif.data <- get(oo)
    for (spp in unique(names(gbif.data$gbif$data))) {
      if ((nrow(gbif.data$gbif$data[[spp]])>0)) {
        xys <- gbif.data$gbif$data[[spp]] %>% dplyr::select(name,latitude,longitude,key) %>% filter(!is.na(longitude)& !is.na(latitude))
        if (nrow(xys)>0) {
          list_spps <- unique(c(list_spps,spp))
          spps <- st_as_sf(xys,coords=c("longitude","latitude"),crs=st_crs(meow)) %>% st_transform(st_crs(mgt.2016.pols))
          mat <- spps %>% st_intersects(mprov.xy,sparse=F) 
          grdXprov %<>% bind_cols(tibble(colSums(mat),.name_repair=~c(spp)))
          mat <- spps %>% st_intersects(spp.grid,sparse=F) 
          grdXcell %<>% bind_cols(tibble(colSums(mat),.name_repair=~c(spp)))
        }
      }
    }
  }
}

grd2Xprov <- tibble(.rows=nrow(mprov.xy))
grd2Xcell <- tibble(.rows=length(raw.grid))

for (spp in list_spps) {
  grd2Xprov %<>% bind_cols(
    grdXprov %>% rowwise() %>% transmute( "{spp}" := sum(c_across(starts_with(spp))))
  )
  grd2Xcell %<>% bind_cols(
    grdXcell %>% rowwise() %>% transmute( "{spp}" := sum(c_across(starts_with(spp))))
  )
  save(file=mi.rda,sppXcell,sppXprov,grd2Xcell,grd2Xprov,mangrove_species,mprov.xy)
   
} 

sppXcell %>% bind_cols(grd2Xcell)

sppXcell %>% bind_cols(grd2Xcell) -> mtzXcell
sppXprov %>% bind_cols(grd2Xprov) -> mtzXprov

save(file=mi.rda,mtzXcell,mtzXprov,mangrove_species,mprov.xy)
 
```

Measure biotic simmilarity between provinces:

```{r}
require(vegan)
ss <- rowSums(mtzXprov)>0

## this looks like a good combination (homogeneous groupings)
d1 <- vegdist(sppXprov[ss,],'kulczynski',binary=T)
h1 <- hclust(d1,method='ward.D2')

lbls <- gsub("West|Western|west|western","W",mprov.xy$PROVINCE)
lbls %<>% gsub("East|Eastern|east|eastern","E",.)
lbls %<>% gsub("North|Northern","N",.)
lbls %<>% gsub("South|Southern","S",.)
lbls %<>% gsub("Central","C",.)
lbls %<>% gsub("Tropical","Trop",.)
lbls %<>% gsub("Temperate","Temp",.)


plot(h1,labels=lbls)
table(cutree(h1,h=.5))
abline(h=.5,lty=2,col=2)

mprov.xy$k_group <- cutree(h1,h=.5) 
mprov.xy %>%  group_by(k_group) %>% summarise -> tst
plot(tst)
```



Cluster analysis for grid cells

```{r}
require(vegan)
ss <- rowSums(sppXcell)>0

## this looks like a good combination (homogeneous groupings)
d1 <- vegdist(sppXcell[ss,],'kulczynski',binary=T)
h1 <- hclust(d1,method='ward.D2')
plot(h1)
table(cutree(h1,k=35))

spp.grid %>% filter(ss) %>% transmute(k_group=cutree(h1,k=35)) %>% plot
spp.grid %>% filter(ss) %>% transmute(k_group=cutree(h1,k=35)) %>% group_by(k_group) %>% summarise -> tst
plot(tst)
```


```{r}
head(res)
table(res$kingdom_name,res$habitat)

res %>% group_by(class_name) %>% summarise(total=n_distinct(scientific_name))

res %>% group_by(order_name) %>% summarise(total=n_distinct(scientific_name))
res %>% group_by(order_name) %>% summarise(total=n_distinct(scientific_name)) %>% arrange(desc(total))
```

Check first how many records are available in GBIF for each species:

```{r}
summary(res$gbif_records)
summary(res$gbif_xy_records)
```



