---
title: "Plot Level 4 units based on marine provinces (MEOW) and mangrove biophysical type data"
editor_options: 
  chunk_output_type: console
---


We intersected the polygon of _Marine ecoregions_ or [MEOW](https://www.worldwildlife.org/publications/marine-ecoregions-of-the-world-a-bioregionalization-of-coastal-and-shelf-areas) with mangrove data from _Global Mangrove Watch_ for the year 2016 (GMW, http://data.unep-wcmc.org/datasets/45) and the _Global biophysical typology of mangroves_ for the year 2016 (https://data.unep-wcmc.org/datasets/48) and we plot the number of distinct groups of mangroves.


```{r}
require(dplyr)
require(magrittr)
require(sf)
require(units)
require(raster)
require(readr)
require(stringr)
require(ggplot2)
require(units)
require(tidyr)
require(scatterpie)
require(ggforce)

source(sprintf("%s/proyectos/IUCN-GET/MFT1.2-intertidal-forest/env/project-env.R",Sys.getenv("HOME")))
setwd(work.dir)
```

## Load data

First we read spatial data from our data sources. Steps for download from the original data sources, and preparation of the data (repairing topological invalid polygons, etc) is documented in the repo https://github.com/unsw-edu-au/cesdata.

We use package `sf` to read the spatial data for the biophysical typology of mangroves:

```{r}

mgv_types <- read_sf(sprintf("%s/ecosystems/global/WCMC-mangrove-types/TNC-BioPhys-Mangrove-valid-output/Mangrove_Typology_v2_2_2016_valid.gpkg",gis.data))

```

And for the intersection with MEOW provinces:

```{r}
mgt.meow.its <- read_sf("intersection-type-meow-provs.gpkg")
```


And extract the data summaries we need:

```{r}

mgt.meow.its %>% mutate(area=st_area(geometry),centroid=st_centroid(geometry)) %>% st_drop_geometry  %>% transmute(ID,PROV_CODE,ECO_CODE, Class, Sedimentar,area,lon=st_coordinates(centroid)[,1],lat=st_coordinates(centroid)[,2]) -> rslts_meow

```

## Summarise data by ecoregion

Now we can summarise by province id and mangrove type:

```{r}

rslts_meow %>% group_by(PROV_CODE,Class,Sedimentar) %>% summarise(total=sum(area) %>% drop_units) %>% pivot_wider(id=c(PROV_CODE), names_from=c(Class,Sedimentar), values_from=total) %>% replace(is.na(.), 0) -> prov_areas

rslts_meow %>% group_by(PROV_CODE) %>% summarise(lon=weighted.mean(lon,drop_units(area)), lat=weighted.mean(lat,drop_units(area))) -> prov_xys

prov_data <- prov_xys %>% full_join(prov_areas)

prov_data %<>% rowwise() %>%
  mutate(radius = sum(c_across(Delta_Terrigenous:Lagoon_Carbonate)))
prov_data %<>% mutate(pradius=log10(radius/1e6))

```


## Plot groups

Check range of coordinates:

```{r}
prov_data %>% summarise(range(lat),range(lon))
```

Background world map:

```{r}

world <- map_data("world")

world.plot <- ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "rosybrown3", fill = "antiquewhite", size = 0.1)  +  theme_void() + coord_fixed(xlim = c(-180, 180), ylim = c(-40, 40))

```

Now we can map by totals for each marine province

```{r}
world.plot + geom_scatterpie(aes(x=lon, y=lat, group=PROV_CODE,r=pradius), data=prov_data, cols=colnames(prov_data)[4:9],color=NA) + labs(title=sprintf("Mangrove aggregated by Marine Provinces: %s distinct units", nrow(prov_data)),caption="Locations: centroid of unit; Size: proportional to log(Area km²)",fill = "Biophysical types") + theme(legend.position = "top",legend.key.size = unit(0.1, "cm")) -> p_provinces # +  geom_scatterpie_legend(prov_data$pradius, x=-140, y=-30,n=3)
p_provinces
```

Export maps using:

```{r,eval=false}

ggsave(plot=p_provinces,file=sprintf('%s/Mangrove-aggregated-by-MEOW-provinces.pdf',work.dir),width=200,height=80,units="mm",device=pdf)
```

Save data to Rdata file:

```{r}


meow %>% filter(PROV_CODE %in% slc) %>% group_by(PROV_CODE,PROVINCE) %>% summarise(geom=st_union(geom)) -> mprovs

save(file=sprintf("%s/www/prov-data.rda",script.dir), prov_table, mprovs)

```


#### Session information:

```{r}
sessionInfo()

```
